---
title: "DATA 608 Homework 1"
author: "Shane Hylton"
date: '2022-09-11'
output:
  html_document:
    toc: TRUE
---

**Principles of Data Visualization and Introduction to ggplot2**

I have provided you with data about the 5,000 fastest growing companies in the US, as compiled by Inc. magazine. lets read this in:

```{r}
inc <- read.csv("https://raw.githubusercontent.com/charleyferrari/CUNY_DATA_608/master/module1/Data/inc5000_data.csv", header= TRUE)

library(tidyverse)
```

And lets preview this data:

```{r}
head(inc)
```
```{r}
summary(inc)
```

Think a bit on what these summaries mean. Use the space below to add some more relevant non-visual exploratory information you think helps you understand this data:

```{r}

inc_state <- inc %>%
  group_by(State) %>% arrange(State) %>% summarize(mean_revenue = mean(Revenue), sd_revenue = sd(Revenue), median_revenue = median(Revenue), iqr_revenue = IQR(Revenue), companies = sum(State == State))


inc_state <- inc_state[order(-inc_state$mean_revenue),]

head(inc_state)


```

### Comments

It is worth noting that median and interquartile range may be better representations of this data, because outliers do not impact the median, while they significantly impact the mean. In Idaho, for example, the mean revenue is the highest of all of the states, but the median revenue ranks $40^{th}$ among all of the states.

When ordered by median revenue, lower populated states like Alaska, Wyoming, and Vermont ranked the highest. 

```{r}

inc_state_median <- inc_state[order(-inc_state$median_revenue),]

head(inc_state_median)

```

I will subset the data to look for the highest median revenue in states with more than 50 companies. 

```{r}

inc_state_sub_50 <- subset(inc_state_median, companies >= 50)

head(inc_state_sub_50)

inc_state_median[which(inc_state_median$companies == max(inc_state_median$companies)), 'State']

```

California has the most companies on the list of the 5000 fastest growing companies. 


## Question 1

Create a graph that shows the distribution of companies in the dataset by State (ie how many are in each state). There are a lot of States, so consider which axis you should use. This visualization is ultimately going to be consumed on a 'portrait' oriented screen (ie taller than wide), which should further guide your layout choices.

```{r}
ggplot(inc, aes(y = State))+
  geom_bar(stat = "count",aes(fill = State))+
  ggtitle("Distribution of Fastest Growing Companies by State")+
  xlab("Count")+
  ylab("State")+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_y_discrete(guide = guide_axis(n.dodge = 2))+
  theme(legend.position = "none")
```

```{r}
ggplot(inc_state_median, aes(x = companies, y = reorder(State,companies)))+
  geom_bar(stat = "identity",aes(fill = State))+
  ggtitle("Distribution of Fastest Growing Companies by State")+
  xlab("Count")+
  ylab("State")+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_y_discrete(guide = guide_axis(n.dodge = 2))+
  theme(legend.position = "none")
```


## Quesiton 2

Lets dig in on the state with the 3rd most companies in the data set. Imagine you work for the state and are interested in how many people are employed by companies in different industries. Create a plot that shows the average and/or median employment by industry for companies in this state (only use cases with full data, use R's `complete.cases()` function.) In addition to this, your graph should show how variable the ranges are, and you should deal with outliers.


#### Comment:


I checked the employee totals, and there were a few outliers above 500 employees that were excessive and would skew the axis, so I limited the y-axis to 500 employees. 


```{r}

inc_state <- inc_state[order(-inc_state$companies),]

inc_state[3,"State"]

ny <- subset(inc,State == "NY")

ny <- ny[complete.cases(ny), ]

ny_ind <- ny %>%
  group_by(Industry) %>% summarize(median_revenue = median(Revenue), median_employees = median(Employees), iqr_employees = IQR(Employees), mean_growth = mean(Growth_Rate))

ggplot(ny, aes(x = Industry, y = Employees))+
  geom_boxplot(outlier.color = "red")+
  ylim(0,500)+
  ggtitle("New York Median Employees by Industry")+
  xlab("Industry")+
  ylab("Employees")+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_x_discrete(guide = guide_axis(n.dodge = 4))+
  theme(axis.text.x = element_text(size = 6))

```

## Question 3

Now imagine you work for an investor and want to see which industries generate the most revenue per employee. Create a chart that makes this information clear. Once again, the distribution per industry should be shown.

I will assume that the investor wants to look nationwide, and not just in NY.

```{r}

inc$rev_emp <- inc$Revenue / inc$Employees

ggplot(inc, aes(x = Industry, y = rev_emp))+
  geom_boxplot(outlier.color = "red")+
  ylim(0,1000000)+
  ggtitle("Median Revenue Per Employee by Industry, Nationwide")+
  xlab("Industry")+
  ylab("Revenue Per Employee")+
  theme(plot.title = element_text(hjust = 0.5))+
  #scale_x_discrete(guide = guide_axis(n.dodge = 4))+
  theme(axis.text.x = element_text(size = 6, angle = 90))

```
